---
title: "wk2-workshop"
author: "Nicholas Lee"
output: 
  github_document:
    toc: true
    toc_depth: 2
    fig_width: 7
    fig_height: 6
date: "2024-01-24"
---

### We have 5798 days of data, starting from Jan 3 2001 to Jan 22 2024.

## 1. Find the following statistics about the S&P returns.

```{r}
df = readRDS("../../../OneDrive - National University of Singapore/Y3S2/DSE3101/data/wk2_stocks.rds")
```
The cumulative returns of the S&P index during this period is `r round(sum(df$SPY_returns) * 100, 2)`%.

The average daily returns of the S&P index during this period is `r round(mean(df$SPY_returns) * 100, 2)`%.

The standard deviation of the daily returns of the S&P index during this period is `r round(sd(df$SPY_returns) * 100, 2)`%.

## 2. Create a plot that shows the prices of S&P in this period.

```{r}
library(ggplot2)
ggplot(data = df, aes(x = date, y = SPY_prices)) +
  geom_line()
```

## 3. Create a plot that shows the total yearly returns of S&P from 2001 to 2023.

```{r message=FALSE}
library(dplyr)
library(lubridate)
df %>%
  mutate(year = year(date)) %>%
  filter(year >= 2001, year <= 2023) %>%
  group_by(year) %>%
  summarise(yearly_returns = sum(SPY_returns)) %>%
  ggplot(aes(x = year, y = yearly_returns)) +
  geom_col()
```

