---
title: "Databases"
output: html_document
date: "2024-02-14"
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dbplyr)
library(DBI)
library(RSQLite)
library(dbplot)
library(bigrquery)
library(Lahman)

# pacman::p_load(tidyverse, #tidyversesuiteofpackages
#                dbplyr, # usedplyrwithdatabase
#                DBI, # databaseinterface
#                RSQLite, # SQLitepackageforR
#                dbplot, # aggregatedataanduseggplot2 withdatabase
#                bigrquery, # GoogleBigQuery inR
#                Lahman) # toydata
```

## Establish database connection to memory
```{r}
# Create and connect to an SQLite database
lite_con = DBI::dbConnect(RSQLite::SQLite(), ":memory:")
```

## Copy database to memory from Lahman library
```{r}
# Copy remote database to local dataframe
copy_to(
 dest = lite_con,
 df = Lahman::People,
 name = "people",
 temporary = FALSE,
 indexes = list("playerID")
)

# Show available tables
dbListTables(lite_con)

# Show available columns
dbListFields(lite_con, "people")
```

## Convert database into tibble
```{r}
# Acts as a tibble, but actually is SQLite data types that works with tibble functions
people_db = tbl(lite_con, "people")
class(people_db)
head(people_db)

# Select columns
people_db %>% select(playerID:birthDay,nameFirst:nameLast)

# Filter based on some condition
people_db %>% filter(birthCountry == "Australia")

# Total number players by birth Country
people_db %>% count(birthCountry, sort = TRUE)

# Mean height & weight for each birthCountry
hw_db = people_db %>%
  # Remove players with unknown birth Country
  filter(!is.na(birthCountry)) %>%
  group_by(birthCountry) %>%
  summarise(mean_weight = mean(weight),
            mean_height = mean(height),
            n = n()) %>%
  arrange(desc(n))

head(hw_db)
```

## Collect data in actual R from database data type
```{r}
# Collect the data back from SQL into R
df_hw = hw_db %>% collect()

class(df_hw) # Proper R tibble, not SQLite data type

# Plot
df_hw %>%
 ggplot(aes(x = mean_weight, y = mean_height, size = n)) +
 geom_point(alpha = 0.5,show.legend = FALSE)
```

## SQL code chunk
```{sql connection = lite_con}
SELECT playerID, birthYear, birthMonth, birthDay, nameFirst, nameLast
FROM People
LIMIT 5
```

```{sql connection = lite_con}
/* Count number of rows*/
SELECT COUNT(*)
FROM people
```

```{sql connection = lite_con}
/* Count distinct values*/
SELECT COUNT(DISTINCT birthCountry)
FROM people
```

```{sql connection = lite_con}
/*  mutate(new_var = var1 * var2) */
SELECT weight, height, (weight/(POWER(height, 2)))*703 AS BMI
FROM people
LIMIT 5
```

## DBI, R code chunk, SQL query inside
```{r}
DBI::dbGetQuery(lite_con, "SELECT playerID, birthYear, birthMonth, birthDay, nameFirst, nameLast
FROM People
LIMIT 5")
```
## Copy new database to memory
```{r}
copy_to(
 dest = lite_con,
 df = Lahman::CollegePlaying,
 name = "collegeplaying",
 temporary = FALSE,
 indexes = list("playerID")
 )

collegeplaying_db = tbl(lite_con,"collegeplaying")

dbListTables(lite_con)

dbListFields(lite_con, "collegeplaying")
```

## SQL code practice
```{sql connection = lite_con}
SELECT people.playerID, nameFirst, nameLast, schoolID, yearID
FROM People
LEFT JOIN collegeplaying
ON people.playerID = collegeplaying.playerID
```

```{sql connection = lite_con}
/* 1. Find all players with first name as “Ken” or “Matty”. */
SELECT nameFirst, nameLast, nameGiven
FROM People
WHERE nameFirst IN ("Ken", "Matty")
```

```{sql connection = lite_con}
/* 2. Find all players whose first names contains “mat”. Then show the unique first names that contain “mat”. */
SELECT DISTINCT nameFirst AS name
FROM People
WHERE nameFirst LIKE "%mat%"
```

```{sql connection = lite_con}
/* 3. Find all players born in Alaska (AK) or Hawaii (HI) after the year of 1980. Order the result in ascending order of birth year. */
SELECT nameFirst, nameLast, birthYear, birthState
FROM People
WHERE (birthState = "AK" OR birthState = "HI") AND (birthYear > 1980)
ORDER BY birthYear
```

```{sql connection = lite_con}
/* 4. Players in the database can attend more than one college. Find the last college attended by each player. */
SELECT people.playerID, nameFirst, nameLast, schoolID, yearID
FROM People
LEFT JOIN CollegePlaying
ON people.playerID = collegeplaying.playerID
GROUP BY people.playerID
HAVING yearID = MAX(yearID)
```

```{sql connection = lite_con}
/* 5. Count the number of unique players who attended Rice University. */
SELECT COUNT(DISTINCT people.playerID)
FROM People
LEFT JOIN CollegePlaying
ON people.playerID = collegeplaying.playerID
WHERE schoolID = "rice"
```

## Disconnect database connection
```{r}
DBI::dbDisconnect(lite_con)
```

## Chinook DB practice
```{r}
lite_con = dbConnect(SQLite(), "../data/chinook.db")
dbListTables(lite_con)
dbListFields(lite_con, "tracks")
```

```{sql connection = lite_con}
/* 1. Find the total number of tracks */
SELECT COUNT(DISTINCT TrackID)
FROM tracks
```
```{sql connection=lite_con}
SELECT * FROM genres
```

```{sql connection = lite_con}
/* 2. Count the number of tracks by each song genre. Sort by descending count, show top 5.*/
SELECT genres.Name, COUNT(DISTINCT TrackID) as n
FROM tracks
LEFT JOIN genres
ON tracks.GenreId = genres.GenreId
GROUP BY tracks.GenreID
ORDER BY n DESC
LIMIT 5
```

```{sql connection = lite_con}
/* 3. Find the occurrences of unique job titles for the employees */
SELECT DISTINCT Title
FROM employees
```

```{r}
# 4. Plot the number of job titles for the employees.
tbl(lite_con, "employees") %>%
  dbplot_bar(Title) +
  labs(title = "Employee Title", y = "Count") +
  theme_classic()
```
## Disconnect DB
```{r}
dbDisconnect(lite_con)
```

## Google BigQuery
```{r}
# Sys.setenv(PROJECT_ID = "____________")

con = dbConnect(
  bigrquery::bigquery(),
  project = "publicdata",
  dataset = "samples",
  billing = Sys.getenv("PROJECT_ID")
)

dbListTables(con)

natality = tbl(con, "natality")
```

```{r}
q1 = "SELECT year, month, day, weight_pounds FROM `publicdata.samples.natality`"
dbGetQuery(con, q1, n = 10)
```

```{r}
df_year = natality %>%
  filter(!is.na(state)) %>%
  group_by(year) %>%
  summarize(total_mil= n()/1e6)

df_year
```

```{r}
# Total newborns over time
natality %>%
  filter(!is.na(state)) %>%
  group_by(year) %>%
  dbplot::dbplot_line(year, total_mil = n()/1e6) +
  theme_minimal() +
  labs(title = "Total new borns (in millions) over time", x = "", y = "")
```

```{r}
# New borns by state
df_state_year = natality %>%
  filter(!is.na(state)) %>%
  group_by(year, state) %>%
  summarize(total_k = n()/1000)

# Select a few states to highlight
df_states = df_state_year %>%
  filter(state %in% c("CA", "TX", "NY", "FL"))

# Plot it
ggplot(data = df_state_year, aes(x = year, y = total_k, group = state)) +
  geom_line(col = "grey75", lwd = 0.5) +
  geom_line(data = df_states, aes(col = state), lwd = 1) +
  scale_color_brewer(palette = "Set1", name = element_blank()) +
  labs(title = "New borns (in thousands) by state over time", x = "", y = "") +
  theme_minimal()
```

## Disconnect database
```{r}
dbDisconnect(con)
```
